---
title: "Data Import"
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
library(haven)
```

### Read in some data

Read in the litters dataset and clean it using clean_names (e.g., turns titles in snakecase)

```{r}
litters_df = read_csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
```

## Take a look at data

Printing in the console, so you can see everything that is going in
```{r}
litters_df
```

Look at head/tail of the litters dataset
```{r}
tail(litters_df)
```

Gives a nice overview of the data
```{r}
skimr::skim(litters_df)
```

View the data in a separate tab using view (dataframe_name) --> but only do that in the console or else the it can mess up the knitting process

## Options to view data

Skip first row , make sure first row is not taken as names , empty values show up as NA

```{r}
litters_df = read_csv("./data/FAS_litters.csv", skip = 10, col_names = FALSE, na = c("", "NA", ".", 999))
```

Check out `?read_csv()` for more information

## Other file formats

Read in an Excel file

```{r}
mlb_df = read_excel("./data/mlb11.xlsx")
mlb_df
```

Read in a SAS file

```{r}
pulse_df = read_sas("./data/public_pulse_data.sas7bdat")
pulse_df
```

## Comparison with Base R

what about`read.csv`

```{r}
litters_base = read.csv("data/FAS_litters.csv")
litters_readr = read_csv("data/FAS_litters.csv")

litters_base
litters_readr
```


## Exporting data

Export the mlb sub-table

```{r}
write_csv(mlb_df, "./data/mlb_subtable.csv")
```






